---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300,
  fig.width = 8
)
```

# fCWTr

<!-- badges: start -->

[![R-CMD-check](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lschneiderbauer/fCWTr/actions/workflows/R-CMD-check.yaml) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/fCWTr)](https://CRAN.R-project.org/package=fCWTr) [![Codecov test coverage](https://codecov.io/gh/lschneiderbauer/fCWTr/branch/master/graph/badge.svg)](https://app.codecov.io/gh/lschneiderbauer/fcwtr?branch=master)

<!-- badges: end -->

The R package fCWTr is a simple wrapper invoking the [fCWT library](https://github.com/fastlib/fCWT), a library implementing a [continuous wavelet transform](https://en.wikipedia.org/wiki/Continuous_wavelet_transform) with a Morlet wavelet, utilizing the power of [fftw](https://www.fftw.org/), a fast fourier transform implementation.

See the original paper by Arts, L.P.A., van den Broek, E.L. The fast continuous wavelet transformation (fCWT) for real-time, high-quality, noise-resistant time–frequency analysis. Nat Comput Sci 2, 47–58 (2022). <https://doi.org/10.1038/s43588-021-00183-z>

## System dependencies

-   R \>= 4.1
-   [fftw](https://www.fftw.org/) library (used by [fCWT](https://github.com/fastlib/fCWT))
-   Optional: a CPU/compiler supporting the [AVX](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set
-   Optional: OpenMP
    -   On Windows, OpenMP support is disabled since rtools decided to compile fftw without OpenMP support.
    -   On Linux and MacOS the build scripts should automatically detect whether OpenMP support is available.

By default, most compiler setups do not make use of AVX to increase portability of the binary. If you are an R user that has a CPU supporting AVX and want to make use of it, you might need to manually enable compiler flags to let R know about it, and install the package from source (so that it gets compiled on your machine). One way to enable the flags is to create a file `~/.R/Makevars` with the following content:

``` bash
CPPFLAGS = -mavx
CXXFLAGS = -mavx
```

## Installation

You can install the latest CRAN release of fCWTr with:

``` r
install.packages("fCWTr")
```

Alternatively, you can install the development version of fCWTr like so (requiring installed [devtools](https://devtools.r-lib.org/) package):

``` r
devtools::install_github("lschneiderbauer/fCWTr")
```

## Example

This is a basic example that invokes the fCWT library to calculate the continuous wavelet transform and plot the result.

```{r example}
library(fCWTr)

# A signal encoded in a numeric vector.
# In this example we use some superimposed sin signals.
signal <- ts_sin_superpos

output <-
  fcwt(
    signal,
    sample_freq = u(44.1, "kHz"),
    freq_begin = u(16, "Hz"),
    freq_end = u(2100, "Hz"),
    n_freqs = 200,
    sigma = 5
  )

# The result is a numeric matrix
dim(output)
```

The result can be easily coerced into a data frame:

```{r example_df}
head(as.data.frame(output), 10)
```

We can also directly plot the resulting scalogram:

```{r example_plot}
plot(output, time_unit = "ms")
```

For long sequences, the required memory can exceed your available local memory. In this case it can be useful to reduce the time resolution of the result and process the data in batches. This can be done with `fcwt_batch()`. In case the batch size is not explicitly provided, some heuristics are used to determine a batch size automatically:

```{r example_long_df, results='hide'}
batch_result <-
  fcwt_batch(
    rep(ts_sin_sin, 5),
    sample_freq = u(44.1, "kHz"),
    freq_begin = u(100, "Hz"),
    freq_end = u(12, "kHz"),
    n_freqs = 200,
    sigma = 4,
    time_resolution = u(10, "ms")
  )

plot(batch_result)
```

<!-- regenerate with devtools::build_readme() -->
