---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fcwtr

<!-- badges: start -->
<!-- badges: end -->

The R package fcwtr is a simple wrapper invoking the [fcwt
library](https://github.com/fastlib/fCWT), a library implementing a
[continuous wavelet transform](https://en.wikipedia.org/wiki/Continuous_wavelet_transform) with a Morlet wavelet, utilizing the power of [fftw](https://www.fftw.org/),
a fast fourier transform implementation.

## Runtime Dependencies
* R >= 4.1 (due to the use of native pipes)
* [fftw](https://www.fftw.org/) library (used by fcwt)


## Installation

You can install the development version of fcwtr like so (requiring installed
[devtools](https://devtools.r-lib.org/) package):

``` r
devtools::install_github("lschneiderbauer/fcwtr")
```

## Example

This is a basic example which shows how to invoke the fcwt library to calculate
the continuous wavelet transform and plot the result.

```{r example}
library(fcwtr)

# You are given some signal encoded in a numeric vector.
# In this example we use some superimposed sin signals.
input <- time_series_sin

# One possibility is to use the fcwt library (almost) directly through
# a thin wrapper:
output <- fcwt(input)

# In this case the result is a tensor of rank 3,
# which can be used for further processing.
# No additional sugar is provided here.
dim(output)

# A little more convenience is provided by `fcwt_df` which returns
# a data frame, already establishes a connection to physical units, and
# allows for pooling, and other small quality-of-live improvements:
df <- fcwt_df(
  time_series_sin,
  sampling_rate = 44100,
  time_resolution = 0.001
)

df

# This data frame can be directly plotted.
plot(df)
```


<!-- regenerate with devtools::build_readme() -->
