---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi=300,
  fig.width=8
)
```

# fcwtr

<!-- badges: start -->
[![R-CMD-check](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R package fcwtr is a simple wrapper invoking the [fcwt
library](https://github.com/fastlib/fCWT), a library implementing a
[continuous wavelet transform](https://en.wikipedia.org/wiki/Continuous_wavelet_transform) with a Morlet wavelet, utilizing the power of [fftw](https://www.fftw.org/),
a fast fourier transform implementation.

## Runtime Dependencies
* R >= 4.1 (due to the use of native pipes)
* [fftw](https://www.fftw.org/) library (used by fcwt)
* A CPU supporting the [AVX2](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set

## Installation

You can install the development version of fcwtr like so (requiring installed
[devtools](https://devtools.r-lib.org/) package):

``` r
devtools::install_github("lschneiderbauer/fcwtr")
```

### Compile-time Dependencies

* Compiler w/ C++17 support
* [fftw](https://www.fftw.org/) library headers
* A CPU/compiler supporting the [AVX2](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set


## Example

This is a basic example which shows how to invoke the fcwt library to calculate
the continuous wavelet transform and plot the result.

```{r example}
library(fcwtr)

# You are given some signal encoded in a numeric vector.
# In this example we use some superimposed sin signals.
input <- ts_sin_superpos

# One possibility is to use the fcwt library (almost) directly through
# a thin wrapper:
output <- fcwt(input)

# In this case the result is a rank 3 tensor,
# which can be used for further processing.
# No additional sugar is provided here.
dim(output)
```

```{r example_df}
# A little more convenience is provided by `fcwt_df` which returns
# a data frame, already establishes a connection to physical units, and
# allows for pooling, and other small quality-of-live improvements:
df <- fcwt_df(
  ts_sin_superpos,
  sampling_rate = 44100,
  time_resolution = 0.001
)

df

# This data frame can be directly plotted.
plot(df)
```

```{r example_long_df, results='hide'}
# For long sequences, the required memory can exceed your local memory. In this
# case it can be useful to reduce the time resolution of the result and process
# the data in batches. This can be done with `fcwt_bulk_df`.
# In case the batch size is not explicitly provided, some heuristics are used to
# determine a batch size automatically:
bulk_df <- fcwt_bulk_df(
  rep(ts_sin_sin, 10),
  sampling_rate = 44100,
  time_resolution = 0.01,
  nsuboctaves = 24L,
  sigma = 4
)

plot(bulk_df)
```

<!-- regenerate with devtools::build_readme() -->
