---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300,
  fig.width = 8
)
```

# fcwtr

<!-- badges: start -->
[![R-CMD-check](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R package fcwtr is a simple wrapper invoking the [fcwt
library](https://github.com/fastlib/fCWT), a library implementing a
[continuous wavelet transform](https://en.wikipedia.org/wiki/Continuous_wavelet_transform) with a Morlet wavelet, utilizing the power of [fftw](https://www.fftw.org/),
a fast fourier transform implementation.

## Runtime Dependencies
* R >= 4.1 (due to the use of native pipes)
* [fftw](https://www.fftw.org/) library (used by fcwt)
* A CPU supporting the [AVX2](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set

## Installation

You can install the development version of fcwtr like so (requiring installed
[devtools](https://devtools.r-lib.org/) package):

``` r
devtools::install_github("lschneiderbauer/fcwtr")
```

### Compile-time Dependencies

* Compiler w/ C++17 support
* [fftw](https://www.fftw.org/) library headers
* A CPU/compiler supporting the [AVX2](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set


## Example

This is a basic example which shows how to invoke the fcwt library to calculate
the continuous wavelet transform and plot the result.

```{r example}
library(fcwtr)

# You are given some signal encoded in a numeric vector.
# In this example we use some superimposed sin signals.
signal <- ts_sin_superpos

# One possibility is to use the fcwt library (almost) directly through
# a thin wrapper:
output <-
  fcwt(
    signal,
    sample_freq = 44100,
    freq_begin = 16,
    freq_end = 2100,
    n_freqs = 200,
    sigma = 5
  )

# In this case the result is a matrix,
# which can be used for further processing.
dim(output)
```

Some helpers to conveniently convert this matrix to a meaningful data frame:

```{r example_df}
as.data.frame(output) |> head(10)
```

We can also directly plot the resulting scalogram:

```{r example_plot}
plot(output)
```


<!-- ```{r example_long_df, results='hide'} -->
<!-- # For long sequences, the required memory can exceed your local memory. In this -->
<!-- # case it can be useful to reduce the time resolution of the result and process -->
<!-- # the data in batches. This can be done with `fcwt_bulk_df`. -->
<!-- # In case the batch size is not explicitly provided, some heuristics are used to -->
<!-- # determine a batch size automatically: -->
<!-- batch_result <- fcwt_batch( -->
<!--   rep(ts_sin_sin, 10), -->
<!--   sampling_rate = 44100, -->
<!--   sigma = 4 -->
<!-- ) -->

<!-- plot(batch_result) -->
<!-- ``` -->

<!-- regenerate with devtools::build_readme() -->
