---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300,
  fig.width = 8
)
```

# fcwtr

<!-- badges: start -->

[![R-CMD-check](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lschneiderbauer/fcwtr/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The R package fcwtr is a simple wrapper invoking the [fcwt library](https://github.com/fastlib/fCWT), a library implementing a [continuous wavelet transform](https://en.wikipedia.org/wiki/Continuous_wavelet_transform) with a Morlet wavelet, utilizing the power of [fftw](https://www.fftw.org/), a fast fourier transform implementation.

## Dependencies

-   R \>= 4.1
-   [fftw](https://www.fftw.org/) library (used by fcwt)
-   Optional: a CPU/compiler supporting the [AVX2](https://en.wikipedia.org/wiki/Advanced_Vector_Extensions) instruction set

If you are an R user that has a CPU supporting AVX2 and want to make use of it, you might need to manually enable compiler flags to let R know about it, and install the package by source (so that it gets compiled on your machine). One way to enable the flags is to create a file `~/.R/Makevars` with the following content:

``` bash
CXXFLAGS = -mavx2
```

## Installation

You can install the development version of fcwtr like so (requiring installed [devtools](https://devtools.r-lib.org/) package):

``` r
devtools::install_github("lschneiderbauer/fcwtr")
```

## Example

This is a basic example that invokes the fcwt library to calculate the continuous wavelet transform and plot the result.

```{r example}
library(fcwtr)

# You are given some signal encoded in a numeric vector.
# In this example we use some superimposed sin signals.
signal <- ts_sin_superpos

output <-
  fcwt(
    signal,
    sample_freq = 44100,
    freq_begin = 16,
    freq_end = 2100,
    n_freqs = 200,
    sigma = 5
  )

# In this case the result is a matrix,
# which can be used for further processing.
dim(output)
```

Conversion functions to a meaningful data frame is also provided:

```{r example_df}
head(as.data.frame(output), 10)
```

We can also directly plot the resulting scalogram:

```{r example_plot}
plot(output)
```

<!-- ```{r example_long_df, results='hide'} -->

<!-- # For long sequences, the required memory can exceed your local memory. In this -->

<!-- # case it can be useful to reduce the time resolution of the result and process -->

<!-- # the data in batches. This can be done with `fcwt_bulk_df`. -->

<!-- # In case the batch size is not explicitly provided, some heuristics are used to -->

<!-- # determine a batch size automatically: -->

<!-- batch_result <- fcwt_batch( -->

<!--   rep(ts_sin_sin, 10), -->

<!--   sampling_rate = 44100, -->

<!--   sigma = 4 -->

<!-- ) -->

<!-- plot(batch_result) -->

<!-- ``` -->

<!-- regenerate with devtools::build_readme() -->
